---
title: "wealth_and_income"
author: "Ijeamaka Anyene"
date: "2/11/2021"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(patchwork)
```

```{r, message = FALSE}
extrafont::loadfonts()
```

```{r}
lifetime_earn <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/lifetime_earn.csv')

student_debt <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/student_debt.csv')

```
Coding Started: 4:37 PM
Ideating Started: 4:48 PM
Coding Started: 4:51 PM
Codong Stopped: 5:01 PM

*SWITCHED IDEAS*
Re-Ideating Started: 5:07 PM
Ideating Stopped: 5:14 PM
Figuring out how to map data to idea: 5:14 PM
Figured it out by 5:20 PM
Coding Started: 5:21 PM
Figuring Out Legend Started: 5:27 PM
Paused Figuring Out Legend to walk Waffles: 5:37 PM
Back! Lets think colors lol: 5:55 PM
Paused: 6:03
Back to Colors: 6:09
Back to Legend/Patchwork Shenanigans: 6:11
Paused for Dinner: 6:33
Back to Patchwork fiddling: 7:08
Finished lol: 8:30
Fiddled around with it for 15 more minutes

```{r}
student_debt_black = filter(student_debt, race == "Black") %>%
  arrange(year) %>%
  mutate(pct_scaled = 10 - (loan_debt_pct*10),
         loan_debt_pct_100 = paste0(round(loan_debt_pct*100, 1), "%"))
```

```{r}
student_debt_list = list()

for(i in 1:nrow(student_debt_black)){
  student_debt_list[[i]] = tibble(
    year = student_debt_black$year[i],
    x = c(0, 10, 10, 0),
    y = c(10, 10, rep(student_debt_black$pct_scaled[i], 2)),
    pct_label = student_debt_black$loan_debt_pct_100[i]
  )
}
```

```{r}
main_plot = 
  ggplot(data = bind_rows(student_debt_list)) +
  geom_segment(aes(x = 0, xend = 10,
                   y = 10, yend = 10),
               size = 2,
               color = "#cdcdc0") +
  geom_polygon(aes(x = x, 
                   y = y,
                   group = year),
               fill = "#0c1931") +
  geom_text(aes(x = 5, y = 0, label = year),
            size = 6,
            family = "Oswald",
            fontface = "bold",
            color = "#626d71") +
  geom_text(aes(x = 5, y = 3, label = pct_label),
            size = 3,
            family = "Open Sans",
            fontface = "bold",
            color = "black") +
  facet_wrap(~year, nrow = 2) +
  coord_polar() +
  ylim(0, 10) +
  labs(title =  "Black Student Loan Debt Crisis") +
  theme_void() +
  theme(
    strip.background = element_blank(),
    strip.text = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold", 
                              size = 30, color = "black", family = "Oswald"),
    panel.background = element_rect(fill = "#f4f2f2",
                                    color = NA),
    plot.background = element_rect(fill = "#f4f2f2",
                                   color = NA),
    plot.margin = margin(5, 0, 0, 0)
  ) +
  NULL

main_plot
```

```{r}
# 10, 25, 50, 75
legend_data = tibble(
  type = c(10, 25, 50, 75),
  type_scales = c(1, 2.5, 5, 7.5),
  label = paste0(type, "%"),
  y = 10 - type_scales
)


legend_list = list()


for(i in 1:nrow(legend_data)){
  legend_list[[i]] = tibble(
    label = legend_data$label[i],
    type = legend_data$type[i],
    x = c(0, 10, 10, 0),
    y = c(10, 10, rep(legend_data$y[i], 2))
  )
}
```

```{r}
legend_plot = 
  ggplot(data = bind_rows(legend_list)) +
  geom_segment(aes(x = 0, xend = 10,
                   y = 10, yend = 10),
               size = 2,
               color = "#cdcdc0") +
  geom_polygon(aes(x = x, 
                   y = y,
                   group = type),
               fill = "#0c1931") +
  facet_wrap(~label, nrow = 1) +
  coord_polar() +
  ylim(0, 10) +
  theme_void() +
  theme(strip.text = element_text(face = "bold", color = "#0c1931",
                                  family = "Open Sans"),
        panel.background = element_rect(fill = "#f4f2f2",
                                    color = NA),
        plot.background = element_rect(fill = "#f4f2f2",
                                   color = NA),
        plot.margin = margin(t = 0, r = 0, b = 0, l = 30))

legend_plot
```


```{r}
title_plot = 
  ggplot() +
  geom_text(aes(x  = 0, y = 2, label = "Percentage of Total Debt Held"),
            family = "Oswald",
            fontface = "bold",
            hjust = 0,
            size = 3) +
  geom_text(aes(x = 0, y = 1, label = "The thickness of the inner perimeter of each circle\n represents of the percentage of the total student loan debt\n held by Black families."),
           family = "Open Sans",
           hjust = 0,
           size = 2) +
  ylim(0, 5) +
  scale_x_continuous(limits = c(0, 3), expand = c(0, 0)) +
  labs(caption = "Data by Urban Institue | Viz by @ijeamaka_a") +
  theme_void() +
  theme(panel.background = element_rect(fill = "#f4f2f2",
                                    color = NA),
        plot.background = element_rect(fill = "#f4f2f2",
                                   color = NA),
        plot.margin = margin(0, 0, 0, 0),
        plot.caption = element_text(size = 6, family = "Open Sans"))

title_plot
```

The thickness of the inner perimeter of each circle represents of the total \nstudent loan debt, what percentage is held by Black families.


Top - Bottom: 1 - 20
Left - Right: 1 - 15
```{r}
layout = c(
  area(t = 1, l = 1, b = 18, r = 15),
  area(t = 19, l = 1, b = 20, r = 5),
  area(t = 16, l = 6, b = 20, r = 15)
)

```



```{r}
main_plot + legend_plot + title_plot + 
  plot_layout(design = layout) 

```


```{r}
ggsave("2021_05_wealth_income.pdf", plot = last_plot(), 
       device = cairo_pdf, path = here::here("outputs"),
       width = 9, height = 6) 


```

